---
title       : Models for Counts
subtitle    : Poisson, Negative Binomial, Zero-Inflated
author      : William Reed
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : mathjax            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
github:
user: strandvej
repo: count



---
## The Observed Value Approach for MNL

> - Fit the Model to the Data
> - Obtain the coefficients and variance covariance matrix.
> - Simulate coefficients from multivariate normal distribution.
> - Use observed values to calculate the probability of being in eaching 
category.
--- .class #id 

---
```{r, echo=FALSE, eval=TRUE} 
require(foreign)
require(mlogit)
ml <- read.dta("http://www.ats.ucla.edu/stat/data/hsbdemo.dta")
ml_1<-mlogit.data(ml,shape="wide",choice="prog")
test <- mlogit(prog ~ 1 | ses + write,  data = ml_1, 
               reflevel="academic", probit=FALSE)
```
--- 


---
Estimated Coefficients
```{r}
test
```
---


---
What can we extract from the text object?

```{r}
coef(test)
```
---

---
What can we extract from the text object?

```{r}
vcov(test)
```
---

---
So Let's do the simulations:

```{r}
#observed values for multinomial
library(MASS)
library(mvtnorm)
n.draws <- 50
set.seed(123)
sim.coefs <- rmvnorm(n.draws, coef(test), vcov(test)) 
head(sim.coefs)
```
---


---
Use the coefficients to caculated probability of being in General or 
Vocational tracks.
```{r,eval=FALSE}
  # second, for current set of coefs, loop through each observation
  # and values for the linear predictors
  for(j in 1:n.obs){
    
    ###reordered columns are now numbered consitently here ###
    Xb.gen[j] <- sim.coefs[i,1] + sim.coefs[i,3] + sim.coefs[i,5]
    Xb.voc[j] <- sim.coefs[i,2] + sim.coefs[i,4] + sim.coefs[i,6]
  }
  ####calculate probability of being in category voc or gen for each observation####
  p.gen=exp(Xb.gen)/(1+exp(Xb.gen)+exp(Xb.voc))
  p.voc=exp(Xb.voc)/(1+exp(Xb.gen)+exp(Xb.voc))
  ####average probability across all observations####
  p.gen[i]<-mean(p.gen)
  p.voc[i]<-mean(p.voc)
}
```
---



---
```{r}
test <- mlogit(prog ~ 1 | ses + write,  data = ml_1, 
               reflevel="academic", probit=FALSE)
test1 <- mlogit(prog ~ 1 | ses + write,  data = ml_1, 
               reflevel="academic", probit=FALSE,alt.subset=c("general","academic"))
hmftest(test,test1)
```
---



---
## Models for Counting: Poisson 
```{r}
count<-rpois(1000,5)
hist(count,col="blue")
```
---

---
## The Poisson Regression Model
> - $E(y)=\mu$
> - $Var(y)=E(y)=\mu$
> - $Pr(y|\mu)=\frac{exp(-\mu)\mu^y}{y!}$
> - $\mu_i=E(y_i|x_i)=exp(\mathbf{x_i\beta})$


---
## R Example:

```{r}
p <- read.csv("http://www.ats.ucla.edu/stat/data/poisson_sim.csv")
p <- within(p, {
  prog <- factor(prog, levels=1:3, labels=c("General", "Academic", "Vocational"))
  id <- factor(id)
})
summary(p)
```
---


---
Describe the Count of Awards
```{r}
library(ggplot2)
ggplot(p, aes(num_awards, fill = prog)) + geom_histogram(binwidth=.5, position="dodge")
```
---

---
```{r}
summary(m1 <- glm(num_awards ~ prog + math, family="poisson", data=p))
```
---


---
```{r}
## calculate and store predicted values
p$phat <- predict(m1, type="response")

## order by program and then by math
p <- p[with(p, order(prog, math)), ]
```
---

---
```{r}
## create the plot
ggplot(p, aes(x = math, y = phat, colour = prog)) +
  geom_point(aes(y = num_awards), alpha=.5, position=position_jitter(h=.2)) +
  geom_line(size = 1) + labs(x = "Math Score", y = "Expected number of awards")
```
---


---

Observed Value Approach for Counts
```{r}
library(obsval)

mod <- obsval(num_awards ~ prog + math, data=p,
              reg.model = "poisson",
              n.draws = 1000,
              effect.var = "math",
              effect.vals = c(65,75), # lowest to mid
              verbose = TRUE)


```
---


---
```{r}
library(vioplot)

```
---


---
```{r}
vioplot(mod$sim.coef[,2], mod$sim.coef[,3], mod$sim.coef[,4], names=c("Academic","Vocational","Math"), 
        col="red")
title("Coefficients")
```
---

---
```{r}
#head(mod$preds)
vioplot(mod$preds[,1], mod$preds[,2], names=c("Low Math","Hight Math"), 
        col="red")
title("Predicted Count")
```
---



---
```{r}
hist(mod$preds[,1],breaks=20,col="red")
hist(mod$preds[,2], breaks=20,col="grey",add=T)
box()
```
---

---
Test of Over dispersion
```{r}
library(AER)
```
---


---
Cameron & Trivedi (1990).
Assume the mean is $E(Y)=\mu$ and the variance is $Var(Y)=\mu$. On may test this assumption as a null hypothesis against an alternative where $Var(Y)=\mu+c∗f(\mu)$ where the constant $c<0$ means under dispersion and $c>0$ means over dispersion. 

$H_0:c=0$ vs. $H_1:c\ne 0$ and the test statistic is asymptotically standard normal under the null.
```{r}
dispersiontest(m1,trafo=1)
```
---


---
```{r}
summary(m2 <- glm.nb(num_awards ~ prog + math, data=p))
```
---

---
```{r,eval=FALSE}

summary(m3 <- zeroinfl(num_awards ~ prog+math| math, data=p))
``` 
---



